---
title: "Practical Machine Learning"

---

###Objective
The objective is that given the input data try to create a model which can predict the classe variable.

```{r}
library(ggplot2)
library(data.table)
library(caret)
```

###Data retrieval and splitting

Load data and split training data into two: 60% for training and the rest for validation.

```{r results="hide"}

train.raw = fread("pml-training-1.csv")
test.raw = fread("pml-testing.csv")

# turn response variable into factor variable
train.raw[,classe:=as.factor(classe)]

idxTrain = createDataPartition(train.raw$classe, p=0.6)[[1]]
train=train.raw[idxTrain,]
validation = train.raw[-idxTrain,]

```

###Data cleaning

```{r}

# remove first seven columns which don't make sense as input
train = train[,!(1:7),with=F]

# remove columns with mostly nan-values
naCols = train[,lapply(train,function(x) sum(is.na(x))/.N)>0.05]

train = train[,!naCols,with=F]

# exclude near zero variance columns
zeroVarCols = nearZeroVar(train)

train = train[,!zeroVarCols,with=F]
validation = validation[,names(train),with=F]

```

###Model estimation

I've decided to try linear discriminant analysis and random forests.Use 5-fold crossvalidation in both cases.

#####Linear discriminant analysis
```{r cache=TRUE}

lda.model <- train(classe ~ ., data=as.data.frame(train), method="lda",trControl=trainControl(method="cv", 5))
lda.model
```

#####Random Forest
```{r cache=TRUE}

rf.model <- train(classe ~ ., data=as.data.frame(train), method="rf",trControl=trainControl(method="cv", 5))
rf.model
```

###Evaluation and model selection
```{r}

confusionMatrix(predict(lda.model,validation),validation[,classe])

confusionMatrix(predict(rf.model,validation),validation[,classe])
```

The random forest model seem to make a much better job than the linear discriminant analysis with an accuracy of 99.8 vs 70.4%. Therefore I select the random forest model for making predictions on the test sample.

####Retrain the selected model on the entire dataset
```{r cache=TRUE}

allData = train.raw[,names(train),with=F]
final.model <- train(classe ~ ., data=as.data.frame(allData), method="rf",trControl=trainControl(method="cv", 5))
final.model
```

###Make test set predictions
``` {r}
cols = head(names(train),-1)
test = test.raw[,cols,with=F]
predictions = predict(final.model,test)

```
